---
title: "Jessica's_class_dplyr_presentation"
author: "Roe"
date: "4/17/2020"
output: html_document
---

tidyverse: dplyr, ggplot2, tidyr, stringr, readr, tibble, purr
"An opinionated collection of R packages designed for data science. All packages share an underlying design philosophy, grammar, and data structure." 

"underlying design philosophy, grammar, and data structure." - this part is important

Look at the data wrangling cheat sheet: https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf

load packages and data
```{r}
library(dplyr)
library(tidyr)
library(titanic)

NR_data <- read.csv("example_data.csv")
```

look at titanic dataset
```{r}
data("titanic_train")

View(titanic_train)
glimpse(titanic_train)
str(titanic_train)
```

***5 verbs of dplyr: select, filter, arrange, mutate, and summarize***

select:
```{r}
select(titanic_train, Fare:Embarked)

titanic_train %>% select(Fare)
titanic_train %>% select(c("Fare", "Ticket", "Embarked"))
```

select with helper functions:
```{r}
# remove certain columns
titanic_train %>% select(-Fare)
titanic_train %>% select(-c("Fare", "Sex", "Age"))

# change the order of columns
titanic_train %>% select(Fare, Sex, Age, everything())

# column name contains a character string
titanic_train %>% select(contains("ar"))

colnames(iris)
iris %>% select(contains("Sepal"))

titanic_train %>% select_if(is.numeric) %>% log()
log(titanic_train)
```

filter:
```{r}
dim(titanic_train)
nrow(titanic_train)

t = 1
t <- 1
t == 2

titanic_train %>% filter(titanic_train$Age > 20)
titanic_train %>% filter(titanic_train$Age > 20 & titanic_train$Age < 40)

titanic_train %>% filter(titanic_train$Sex == "female")
titanic_train %>% filter(titanic_train$Sex != "female") # != means 'NOT EQUAL TO'

titanic_train %>% filter(titanic_train$Pclass %in% 1:2) # %in% is 'group membership' 
# this line says, filter the rows to those that have a Pclass IN the vector 1:2

titanic_train %>% slice(1:6)
head(titanic_train)
```

arrange
```{r}
titanic_train %>% arrange(Survived)
titanic_train %>% arrange(desc(Survived))

titanic_train %>% arrange(Survived, Pclass)
```

mutate
```{r}
titanic_train %>% mutate(log_fare = log(Fare))
```

summarise
```{r}
iris

iris %>% group_by(Species) %>% summarise(n())

iris %>% group_by(Species) %>% summarise(Individuals = n(), Average_Sepal_Length = mean(Sepal.Length)) %>% arrange(desc(Average_Sepal_Length))
```


```{r}
View(NR_data)

# Change improperly named column name
colnames(NR_data)[which(names(NR_data) == "NE")] <- "Subplot"

NR_data <- read.csv("example_data.csv")

NR_data <- NR_data %>% rename(Subplot = NE)

# Reduce data frame
NR_data_reduced <- NR_data %>% select(Plot, Subplot, Species, LAI)
NR_data_reduced

# Back to summarising
NR_data_reduced2 <- NR_data_reduced %>% group_by(Plot, Species) %>% summarise(LAI = sum(LAI/5))

NR_data_reduced2

# Spread data to wide format
wide_NR_data_reduced2 <- NR_data_reduced2 %>% spread(Species, LAI)
wide_NR_data_reduced2

# Change plot column to rownames
wide_NR_data_reduced2 %>% column_to_rownames("Plot")

# With piping, we can avoid creating all these new variables though
NR_data_reduced <- NR_data %>% select(Plot, Subplot, Species, LAI) %>% group_by(Plot, Species) %>% summarise(LAI = sum(LAI/5)) %>% spread(Species, LAI) %>% column_to_rownames("Plot")

View(NR_data_reduced)



```

